---
title: "Assignment 2"
author: "Dhwani Kirankumar Trivedi, Ankita Ajit Dodhihal, Miraj Bhimani"
date: "9/25/2021"
output: html_document
---

```{r}
#Question 1 Generate a vector  x<-1:1000
user <-  function(n){
  x <- c(1:n)
}
x <- user(1000)
x
```

```{r}
#Question 2
x = 1:10
length(x)
for (i in 1:length(x)){
  print(i)
}
```

```{r}
#Question 3
x =  c(1:10)
for (i in x[]){
  i = i*2
  print(i)
}
```

```{r}
#Question 4
x =  c(1:10)
t1 = Sys.time()
for (i in x[]){
  i = i*2
  print(i)
}
t2 <- Sys.time()
ct <- t2-t1
print(paste("The computation time is " ,ct))
```
```{r}
#Question 5
user <-  function(n){
  x = c(1:n)
  x <- x*2
  
}
t3 <- Sys.time()
user(10)
t4 <- Sys.time()
t4-t3


```

```{r}
#Question 7 
#For x <- 1:100000
t3 <- Sys.time()
user(100000)
t4 <- Sys.time()
t4-t3

```


---
#Question 1.8 Solved by Ankita Dodhihal Dhwani Trivedi Miraj Bhimani

#The following observations have been recorded for vector length 
#1. The computation time for the vector length 10000 is 0.008435965 secs.
#2. The computation time for the vector length 100000 is 0.02526903 secs. 
#3. The computation time for the vector length 1000000 is 0.02483988 secs. 

It can be clearly observed that the the effect of vector length is directly proportional to the computation time. 
It also represents that the CPU time will be linearly grow with respect to the vector length.
representation of the computation time can be done using Big O Notation as well i.e. O(1)

#Effect of for loop on computation time.

It can be observed that the for loop iterates n times. 
The computation time will be T(n) = O(n) n nano seconds and the CPU time will grow linearly with respect to program variable n.
---


```{r}
#For x 100000

t3 <- Sys.time()
user(100000)
t4 <- Sys.time()
t4-t3
```

```{r}
t3 <- Sys.time()
user(100000)
t4 <- Sys.time()
c<-t4-t3

```

```{r}
#Question 2.1
# Solved by Ankita Dodihal Dhwani Trivedi Miraj Bhimani
library(dplyr)
rows <- 1:1000
cols <- 1:10000
#Generate a random dataframe
df <- data.frame(rows,cols)
df

frac_rows <- function(df, n){
  if(is.numeric(n)){
  print("The sample is")
  }
  else{
    stop("The fraction is not an integer")
  }
  c <- sample_n(df,n)
return (c)
}

frac_rows(df, 1000)
```

```{r}
#3.1 Solved by Ankita Dodihal Dhwani Trivedi Miraj Bhimani

df<-read_csv("~/Downloads/COVID_19_Nursing_Home_Data_2021_09_12-2.csv")

ans3.1<-df %>%
  group_by(provider_state) %>%
  summarise(count=(length(unique(federal_provider_number)))) %>% 
  arrange(desc(count))
ans3.1
```


```{r}
#3.2 Solved by Ankita Dodihal Dhwani Trivedi Miraj Bhimani
county_df<-df %>%
  group_by(county) %>%
  summarise(count=(length(unique(federal_provider_number)))) %>% 
  
  arrange(desc(count)) %>%
  slice(1:5)
county_df

```


```{r}
#Question 3.3 Solved by Ankita Dodihal Dhwani Trivedi Miraj Bhimani

df <- read.csv("~/Downloads/COVID_19_Nursing_Home_Data_2021_09_12-2.csv")
df
#The shortage of Nursing staff here corresponds to the no of hospitals having shortage by the week in a particular state however there might be some hospitals which did not have any shortage hence, the approach is to count the number of shortage of nursing staff in the state by federal provider number. For example x hospital in y state did not have shortage in nursing staff hence we will filter the shortage values accordingly. 

total<- df%>% 
  group_by(provider_state)%>%
  summarise(count = (length(unique(federal_provider_number))))
total


```

```{r}
#Table 1
table_1<- distinct(select(df,week_ending,provider_state,federal_provider_number,shortage_of_nursing_staff))%>%
  filter(shortage_of_nursing_staff == "Y")%>%
  arrange(week_ending, provider_state)%>%
  group_by(provider_state, federal_provider_number)%>%
  summarise(weeks=n())
table_1

temp_1<- table_1%>% 
  group_by(provider_state)%>%
  summarise(no_of_providers=n())
temp_1

temp_2 <- table_1%>%
  filter(weeks>20)%>%
  group_by(provider_state)%>%
  summarise(providers_more_than_20=n())
temp_2

#Merge rows with different columns
merge_table <- merge(total, temp_2, by = "provider_state", all = T)%>% drop_na()
merge_table


merge_table$percentage<- ((merge_table[,3]/merge_table[,2])*100)
merge_table

#To calculate the threshold value
th_1 <- mean(merge_table$percentage)
th_1

th <- 10
#Approach 1 To extract the state using mean value
ans_1 <- merge_table%>%
  filter(percentage>th)

ans_1<- select(ans_1,provider_state)
ans_1

#Approach 2 to calculate the mean values by calling the threshold function
threshold_1 <- function(th){
  merge_table%>%
  filter(percentage>th)
  
}
th_10 <- threshold_1(10)
th_20 <- threshold_1(20)
th_30 <- threshold_1(30)


```

```{r}
#Table 2
table_2<- distinct(select(df,week_ending,provider_state,federal_provider_number,shortage_of_clinical_staff))%>%
  filter(shortage_of_clinical_staff == "Y")%>%
  arrange(week_ending, provider_state)%>%
  group_by(provider_state, federal_provider_number)%>%
  summarise(weeks=n())
table_2

#Calculating total no of providers who had shortage of clinical staff
total_providers<- table_2%>% 
  group_by(provider_state)%>%
  summarise(no_of_providers=n())
total_providers

#Calculating total no of providers who had shortage of clinical staff of more than 20 weeks
providers_more_than_20 <- table_2%>%
  filter(weeks>20)%>%
  group_by(provider_state)%>%
  summarise(no_of_providers_more_than_20=n())
providers_more_than_20

#Merging two tables with different dimensions 
merge_tables <- merge(total, providers_more_than_20, by = "provider_state", all = T)%>% drop_na()
merge_tables

merge_tables$percentage<- ((merge_tables[,3]/merge_tables[,2])*100)
merge_tables

th <- mean(merge_tables$percentage)
th

#Approach A
ans2 <- merge_tables%>%
  filter(percentage>th)
ans2<- select(ans2, provider_state)
ans2


#Approach B
threshold_1 <- function(th){
  merge_tables%>%
  filter(percentage>th)
}
th_10 <- threshold_1(10)
th_20 <- threshold_1(20)
th_30 <- threshold_1(30)
th_40 <- threshold_1(40)
```

```{r}
#Table 3 
#shortage_of_aides
table_3<- distinct(select(df,week_ending,provider_state,federal_provider_number,shortage_of_aides))%>%
  filter(shortage_of_aides == "Y")%>%
  arrange(week_ending, provider_state)%>%
  group_by(provider_state, federal_provider_number)%>%
  summarise(providers_with_shortage=n())
table_3

#Calculating total no of providers who had shortage of aides
total_providers<- table_3%>% 
  group_by(provider_state)%>%
  summarise(no_of_providers=n())
total_providers

#Calculating total no of providers who had shortage of aides of more than 20 weeks
providers_more_than_20 <- table_3%>%
  filter(providers_with_shortage>20)%>%
  group_by(provider_state)%>%
  summarise(no_of_providers_more_than_20=n())
providers_more_than_20

#Merging two tables with different dimensions 
merge_tables <- merge(total, providers_more_than_20, by = "provider_state", all = T)%>% drop_na()
merge_tables

merge_tables$percentage<- ((merge_tables[,3]/merge_tables[,2])*100)
merge_tables

th <- mean(merge_tables$percentage)
th
#Approach A
ans3 <- merge_tables%>%
  filter(percentage>th)
ans3<- select(ans3, provider_state)
ans3

threshold_1 <- function(th){
  merge_tables%>%
  filter(percentage>th)
}
th_10 <- threshold_1(10)
th_20 <- threshold_1(20)
th_30 <- threshold_1(30)
th_40 <- threshold_1(40)
```

```{r}
#Table 4
#shortage_of_other_staff
library(tidyr)
library(dplyr)
table_4<- distinct(select(df,week_ending,provider_state,federal_provider_number,shortage_of_other_staff))%>%
  filter(shortage_of_other_staff == "Y")%>%
  arrange(week_ending, provider_state)%>%
  group_by(provider_state, federal_provider_number)%>%
  summarise(weeks=n())
table_4
#Calculating total no of providers who had shortage of clinical staff
total_providers<- table_4%>% 
  group_by(provider_state)%>%
  summarise(no_of_providers=n())
total_providers

#Calculating total no of providers who had shortage of clinical staff of more than 20 weeks
providers_more_than_20 <- table_4%>%
  filter(weeks>20)%>%
  group_by(provider_state)%>%
  summarise(no_of_providers_more_than_20=n())
providers_more_than_20

#Merging two tables with different dimensions 
merge_tables <- merge(total, providers_more_than_20, by = "provider_state", all = T)%>% drop_na()
merge_tables

#To calculate the percentage of providers who had shortage of aides more than 20 weeks
merge_tables$percentage<- ((merge_tables[,3]/merge_tables[,2])*100)
merge_tables

#library(tidyverse)
#in3ca<-cbind(temp_3c_a1,temp_3c_a2)
#fin3ca<- fin3ca[,-3]
#fin3ca$threshold<- (fin3ca[,3] / fin3ca[,2])*100
#fin3ca

# Approach A (Calculating the threshold value by calculating mean)
th_mean <- mean(merge_tables$percentage)
th_mean

merge_tables%>%
  filter(percentage>th_mean)
merge_tables$provider_state

#Approach B Set the threshold values by creating a function
threshold_1 <- function(th){
  merge_tables%>%
  filter(percentage>th)
}
th_10 <- threshold_1(10)
th_20 <- threshold_1(20)
th_30 <- threshold_1(30)
th_40 <- threshold_1(40)


```

```{r}
#Question 3.4 Solved by Ankita Dodhihal Dhwani Trivedi Miraj Bhimani 

#Table1

library(dplyr)
library(tidyr)
#Creating a data frame with the county name, federal provider number and the weeks of shortage of N95 masks for that corresponding provider in the county.

table_4_1<- distinct(select(df,week_ending,county,federal_provider_number,any_current_supply_of_n95_masks))%>%
  filter(any_current_supply_of_n95_masks == "N")%>%
  arrange(week_ending, county)%>%
  group_by(county, federal_provider_number)%>%
  summarise(weeks=n())
table_4_1

#Creating a data frame with the total number of unique providers in a county
temp1_t1<- df%>% 
  group_by(county)%>%
  summarise(total_providers=(length(unique(federal_provider_number))))
temp1_t1

#Creating a data frame with the total number of providers in a county that experienced a shortage for more than 10 weeks
temp2_t1 <- table_4_1%>%
  filter(weeks>10)%>%
  group_by(county)%>%
  summarise(providers_more_than_10=n())
temp2_t1

#Merge rows with different columns - Merged the above data frames and dropped the counties that did not face a shortage of more than 10 weeks(i.e the NA values)
merge_table41 <- merge(temp1_t1, temp2_t1, by = "county", all = T)%>% drop_na()
merge_table41

#Creating a calculated column to find the percentage of providers that faced over 10 weeks of shortage by the total providers that faced a shortage 
merge_table41$percentage<- ((merge_table41[,3]/merge_table41[,2])*100)
merge_table41

#To calculate the threshold value
th_4_t1 <- mean(merge_table41$percentage)
th_4_t1

#Counties that had shortage more than our threshold value
ans3.4_1 <- merge_table41%>%
  filter(percentage>th_4_t1)
ans3.4_1<- select(ans3.4_1,county)
ans3.4_1

# Approach B to calculate the number of couties which have more than 10, 20, 30, 40
threshold_1 <- function(th){
  merge_table41%>%
  filter(percentage>th)
}
th_1 <- threshold_1(10)
th_1
th_2 <- threshold_1(20)
th_2
th_3 <- threshold_1(30)
th_3
th_4 <- threshold_1(90)
th_4
```

```{r}
#3.4 
#Table2

#Creating a data frame with the county name, federal provider number and the weeks of shortage of surgical masks for that corresponding provider in the county.

table_4_2<- distinct(select(df,week_ending,county,federal_provider_number,any_current_supply_of_surgical_masks))%>%
  filter(any_current_supply_of_surgical_masks == "N")%>%
  arrange(week_ending, county)%>%
  group_by(county, federal_provider_number)%>%
  summarise(weeks=n())
table_4_2

#Creating a data frame with the total number of unique providers in a county
temp1_t1<- df%>% 
  group_by(county)%>%
  summarise(total_providers=(length(unique(federal_provider_number))))
temp1_t1

#Creating a data frame with the total number of providers in a county that experienced a shortage for more than 10 weeks
temp2_t2 <- table_4_2%>%
  filter(weeks>10)%>%
  group_by(county)%>%
  summarise(providers_more_than_10=n())
temp2_t2

#Merge rows with different columns - Merged the above data frames and dropped the counties that did not face a shortage of more than 10 weeks(i.e the NA values)
merge_table42 <- merge(temp1_t1, temp2_t2, by = "county", all = T)%>% drop_na()
merge_table42

#Creating a calculated column to find the percentage of providers that faced over 10 weeks of shortage by the total providers in the county 
merge_table42$percentage<- ((merge_table42[,3]/merge_table42[,2])*100)
merge_table42

#To calculate the threshold value
th_4_t2 <- mean(merge_table42$percentage)
th_4_t2

#Counties that had shortage more than our threshold value
ans3.4_2 <- merge_table42%>%
  filter(percentage>th_4_t2)
ans3.4_2

ans3.4_2<- select(ans3.4_2,county)
ans3.4_2

# Approach B to calculate the number of couties which have more than 10, 20, 30, 40
threshold_1 <- function(th){
  merge_table42%>%
  filter(percentage>th)
}
th_1 <- threshold_1(10)
th_1
th_2 <- threshold_1(20)
th_2
th_3 <- threshold_1(30)
th_3
th_4 <- threshold_1(40)
th_4

```

```{r}
#3.4 
#Table3

library(dplyr)
library(tidyr)
#Creating a data frame with the county name, federal provider number and the weeks of shortage of eye protection for that corresponding provider in the county.

table_4_3<- distinct(select(df,week_ending,county,federal_provider_number,any_current_supply_of_eye_protection))%>%
  filter(any_current_supply_of_eye_protection == "N")%>%
  arrange(week_ending, county)%>%
  group_by(county, federal_provider_number)%>%
  summarise(weeks=n())
table_4_3

#Creating a data frame with the total number of unique providers in a county
temp1_t1<- df%>% 
  group_by(county)%>%
  summarise(total_providers=(length(unique(federal_provider_number))))
temp1_t1

#Creating a data frame with the total number of providers in a county that experienced a shortage for more than 10 weeks
temp2_t3 <- table_4_3%>%
  filter(weeks>10)%>%
  group_by(county)%>%
  summarise(providers_more_than_10=n())
temp2_t3

#Merge rows with different columns - Merged the above data frames and dropped the counties that did not face a shortage of more than 10 weeks(i.e the NA values)
merge_table43 <- merge(temp1_t1, temp2_t3, by = "county", all = T)%>% drop_na()
merge_table43

#Creating a calculated column to find the percentage of providers that faced over 10 weeks of shortage by the total providers in the county 
merge_table43$percentage<- ((merge_table43[,3]/merge_table43[,2])*100)
merge_table43

#To calculate the threshold value
th_4_t3 <- mean(merge_table43$percentage)
th_4_t3

#Counties that had shortage more than our threshold value
ans3.4_3 <- merge_table43%>%
  filter(percentage>th_4_t3)
ans3.4_3

ans3.4_3<- select(ans3.4_3,county)
ans3.4_3

# Approach B to calculate the number of couties which have more than 10, 20, 30, 40
threshold_1 <- function(th){
  merge_table43%>%
  filter(percentage>th)
}
th_1 <- threshold_1(10)
th_1
th_2 <- threshold_1(20)
th_2
th_3 <- threshold_1(30)
th_3
th_4 <- threshold_1(9)
th_4
```

```{r}
#3.4
#Table4

library(dplyr)
library(tidyr)
#Creating a data frame with the county name, federal provider number and the weeks of shortage of gowns for that corresponding provider in the county.

table_4_4<- distinct(select(df,week_ending,county,federal_provider_number,any_current_supply_of_gowns))%>%
  filter(any_current_supply_of_gowns == "N")%>%
  arrange(week_ending, county)%>%
  group_by(county, federal_provider_number)%>%
  summarise(weeks=n())
table_4_4

#Creating a data frame with the total number of unique providers in a county
temp1_t1<- df%>% 
  group_by(county)%>%
  summarise(total_providers=(length(unique(federal_provider_number))))
temp1_t1

#Creating a data frame with the total number of providers in a county that experienced a shortage for more than 10 weeks
temp2_t4 <- table_4_4%>%
  filter(weeks>10)%>%
  group_by(county)%>%
  summarise(providers_more_than_10=n())
temp2_t4

#Merge rows with different columns - Merged the above data frames and dropped the counties that did not face a shortage of more than 10 weeks(i.e the NA values)
merge_table44 <- merge(temp1_t1, temp2_t4, by = "county", all = T)%>% drop_na()
merge_table44

#Creating a calculated column to find the percentage of providers that faced over 10 weeks of shortage by the total providers in the county 
merge_table44$percentage<- ((merge_table44[,3]/merge_table44[,2])*100)
merge_table44

#To calculate the threshold value
th_4_t4 <- mean(merge_table44$percentage)
th_4_t4

#Counties that had shortage more than our threshold value
ans3.4_4 <- merge_table44%>%
  filter(percentage>th_4_t4)
ans3.4_4

ans3.4_4<- select(ans3.4_4,county)
ans3.4_4

# Approach B to calculate the number of couties which have more than 10, 20, 30, 40
threshold_1 <- function(th){
  merge_table44%>%
  filter(percentage>th)
}
th_1 <- threshold_1(10)
th_1
th_2 <- threshold_1(20)
th_2
th_3 <- threshold_1(30)
th_3
th_4 <- threshold_1(90)
th_4
```


```{r}
library(dplyr)
library(tidyr)
#Creating a data frame with the county name, federal provider number and the weeks of shortage of gloves for that corresponding provider in the county.

table_4_5<- distinct(select(df,week_ending,county,federal_provider_number,any_current_supply_of_gloves))%>%
  filter(any_current_supply_of_gloves== "N")%>%
  arrange(week_ending, county)%>%
  group_by(county, federal_provider_number)%>%
  summarise(weeks=n())
table_4_5

#Creating a data frame with the total number of unique providers in a county
temp1_t1<- df%>% 
  group_by(county)%>%
  summarise(total_providers=(length(unique(federal_provider_number))))
temp1_t1

#Creating a data frame with the total number of providers in a county that experienced a shortage for more than 10 weeks
temp2_t5 <- table_4_5%>%
  filter(weeks>10)%>%
  group_by(county)%>%
  summarise(providers_more_than_10=n())
temp2_t5

#Merge rows with different columns - Merged the above data frames and dropped the counties that did not face a shortage of more than 10 weeks(i.e the NA values)
merge_table45 <- merge(temp1_t1, temp2_t5, by = "county", all = T)%>% drop_na()
merge_table45

#Creating a calculated column to find the percentage of providers that faced over 10 weeks of shortage by the total providers in the county 
merge_table45$percentage<- ((merge_table45[,3]/merge_table45[,2])*100)
merge_table45

#To calculate the threshold value
th_4_t5 <- mean(merge_table45$percentage)
th_4_t5

#Counties that had shortage more than our threshold value
ans3.4_5 <- merge_table45%>%
  filter(percentage>th_4_t5)
ans3.4_5

ans3.4_5<- select(ans3.4_5,county)
ans3.4_5

# Approach B to calculate the number of couties which have more than 10, 20, 30, 40
threshold_1 <- function(th){
  merge_table45%>%
  filter(percentage>th)
}
th_1 <- threshold_1(10)
th_1
th_2 <- threshold_1(20)
th_2
th_3 <- threshold_1(30)
th_3
th_4 <- threshold_1(90)
th_4
```

```{r}
#3.4
#Table6

library(dplyr)
library(tidyr)
#Creating a data frame with the county name, federal provider number and the weeks of shortage of sanitizer for that corresponding provider in the county.

table_4_6<- distinct(select(df,week_ending,county,federal_provider_number,any_current_supply_of_hand_sanitizer))%>%
  filter(any_current_supply_of_hand_sanitizer== "N")%>%
  arrange(week_ending, county)%>%
  group_by(county, federal_provider_number)%>%
  summarise(weeks=n())
table_4_6

#Creating a data frame with the total number of unique providers in a county
temp1_t1<- df%>% 
  group_by(county)%>%
  summarise(total_providers=(length(unique(federal_provider_number))))
temp1_t1

#Creating a data frame with the total number of providers in a county that experienced a shortage for more than 10 weeks
temp2_t6 <- table_4_6%>%
  filter(weeks>10)%>%
  group_by(county)%>%
  summarise(providers_more_than_10=n())
temp2_t6

#Merge rows with different columns - Merged the above data frames and dropped the counties that did not face a shortage of more than 10 weeks(i.e the NA values)
merge_table46 <- merge(temp1_t1, temp2_t6, by = "county", all = T)%>% drop_na()
merge_table46

#Creating a calculated column to find the percentage of providers that faced over 10 weeks of shortage by the total providers in the county 
merge_table46$percentage<- ((merge_table46[,3]/merge_table46[,2])*100)
merge_table46

#To calculate the threshold value
th_4_t6 <- mean(merge_table46$percentage)
th_4_t6

#Counties that had shortage more than our threshold value
ans3.4_6 <- merge_table46%>%
  filter(percentage>th_4_t6)
ans3.4_6
ans3.4_6<- select(ans3.4_6,county)
ans3.4_6

# Approach B to calculate the number of couties which have more than 10, 20, 30, 40
threshold_1 <- function(th){
  merge_table46%>%
  filter(percentage>th)
}
th_1 <- threshold_1(10)
th_1
th_2 <- threshold_1(20)
th_2
th_3 <- threshold_1(30)
th_3
th_4 <- threshold_1(90)
th_4
```

```{r}
#3.5
#Creating a data frame with the federal provider name and the weeks of shortage of ventilators for that corresponding provider.
table_5_1<- distinct(select(df,week_ending,county,federal_provider_number,provider_name ,any_current_supply_of_ventilator_supplies))%>%
  filter(any_current_supply_of_ventilator_supplies== "N")%>%
  group_by(provider_name)%>%
  summarise(weeks=n())%>%
  filter(weeks>10) #filtering out provider that faced less than or equal to 10 weeks of shortage
table_5_1

```

---
Contribution: Everyone had equal contribution to the assignment. 
---